---
title: 'Lab 7: Stochastic Compartmental Models and Monte Carlo Intervals'
author: "YOUR NAME HERE"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)

## function to calculate lambda_t
lambda_t_fcn <- function(RN, D, I_i, N){
  c_e <- RN/D
  return(1-exp(-c_e*I_i/N))
}
```

In your previous lab, you wrote a function for a stochastic SIR model, and plotted the results of two realizations of that stochastic process. While two realizations let us visualize how the stochastic nature of these models can lead to different results, in practice, we will want to (1) run more than two simulations to understand the variability in the process of interest, (2) reshape simulation results from wide form to long form more efficiently for plotting, (3) summarize that variability using some kind of confidence interval, and (4) create plots to visualize our results.

# SIR model

The difference equations for the stochastic version of the SIR model are given below:

$$
S_t=S_{t-1}-I_t^{(SI)}\\
I_t=I_{t-1}+I_t^{(SI)}-R_t^{(IR)}\\
R_t=R_{t-1}+R_t^{(IR)}\\
$$

Remember now we have random variables in our difference equations: 

- $I_t^{(SI)}$: number of newly infectious individuals at time t
- $R_t^{(IR)}$: number of newly recovered individuals at time t

We will use the chain binomial structure that we discussed in the previous class, which just means that 

- $I_t^{(SI)}\sim Binomial(S_{t-1}, \lambda_t)$
- $R_t^{(IR)}\sim Binomial(I_{t-1}, \rho)$, where $\rho=1-e^{-1/D}$

Below I have included a version of the function for the stochastic SIR simulation, which should be similar to the function that you wrote in your previous lab.

```{r}
SIR_simulation <- function(N, S0, I0, RN, D, Time, wideform=FALSE){
  SIR_df <- data.frame(time=0:Time,
                     S=rep(NA, Time+1),
                     I=rep(NA, Time+1),
                     R=rep(NA, Time+1),
                     I_SI=rep(NA, Time+1),
                     R_IR=rep(NA, Time+1),
                     lambda_t=rep(NA, Time+1))

  SIR_df$S[1] <- S0
  SIR_df$I[1] <- I0
  SIR_df$R[1] <- N-S0-I0
  
  rho <- 1-exp(-1/D)

  for (t in 2:(Time+1)){
    SIR_df$lambda_t[t] <- lambda_t_fcn(RN=RN, D=D, I_i=SIR_df$I[t-1], N=N)
    SIR_df$I_SI[t] <- rbinom(n=1, size=SIR_df$S[t-1], prob=SIR_df$lambda_t[t])
    SIR_df$R_IR[t] <- rbinom(n=1, size=SIR_df$I[t-1], prob=rho)
    SIR_df$S[t] <- SIR_df$S[t-1]-SIR_df$I_SI[t]
    SIR_df$I[t] <- SIR_df$I[t-1]+SIR_df$I_SI[t]-SIR_df$R_IR[t]
    SIR_df$R[t] <- SIR_df$R[t-1]+SIR_df$R_IR[t]
  }
  
  if (wideform==TRUE){
    return(SIR_df)
  }
  else{
      return(data.frame(time=rep(0:Time, 3),
                    compartment=rep(c("S","I", "R"), each=(Time+1)),
                    count=c(SIR_df$S, SIR_df$I, SIR_df$R)))
  }
}
```

## Wide form/long form

**(a) One difference between the above function and what you wrote is the inclusion of the variable `wideform` as an argument in the function. By default, it is set to FALSE, and in this case should return a data frame with 3 columns that is hard to read, but facilitates plotting in ggplot2. Confirm that this is what it returns by running the code below.**

```{r}
SIR_sim1 <- SIR_simulation(N=1000, S0=999, I0=1, RN=4, D=3, Time=28, wideform=FALSE)
head(SIR_sim1)
```

**(b) What happens when you set `wideform=TRUE` (just copy the code from lines 83-84 and change FALSE to TRUE and rename your object as SIR_sim1_wide). Use `head` to print the first 6 lines - what do they look like?**

```{r}

```

**(c) What happens if you omit `wideform` from the list of arguments all together? Why?**

```{r}

```

You will get to explore other ways of reshaping data in your homework.

## Running and storing multiple realizations (iterations) of the simulation

**(d) We want run multiple simulations, perhaps hundreds. What are some ways we could think about using our existing function to do this? Let's brainstorm together and try some things out.**

```{r}

```

## Monte Carlo confidence intervals

Loosely, Monte Carlo (MC) intervals can be obtained from many simulations, and can be used to obtain confidence intervals. They can be particularly useful in cases where analytic confidence intervals (like those obtained based on a t-distribution, for example), are not available, or when the assumptions for calculating such intervals are not satisfied, making them unreliable. To calculate these intervals, we will need the `quantile` function from R.

**(e) Run the following example to (re)-familiarize yourself with this function.**

```{r}
## returns 2.5 and 97.5 percentiles
quantile(x=1:10, probs=c(0.025, 0.975))
```

**(f) Modify the code in (e) to return the 5th and 95th percentiles.**

```{r}

```

**(g) How can we use this on our simulation results from (d)? Let's brainstorm together and try some things out.**

```{r}

```

## Plotting MC intervals

**(h) Now we want to plot our results, including our MC intervals. We will need to add extra layers to our ggplot to accomplish this.**

```{r}

```

